---
title: "Project"
author: "Trinath Sai Subhash Reddy Pittala, Uma Maheswara R Meleti, Hemanth Vasireddy"
date: "`r Sys.Date()`"
output:
  pdf_document:
    keep_tex: yes
  html_document:
    code_folding: hide
    word_wrap: true
---

```{r setup, include=FALSE}
library(formatR)
library(tidyverse)
library(GGally)
library(ggfortify)
library(nycflights13)
library(modelr)
library(magick)
library(haven)
library(dplyr)
library(ISLR)
library(broom)
library(lubridate)
library(ggplot2)
library(nycflights13)
library(knitr)
library(caret)
library(car)
library(kableExtra)
library(glmnet)
library(corrplot)
knitr::opts_chunk$set(width = 60, warning = FALSE, message = FALSE, tidy.opts=list(width.cutoff=60),tidy=TRUE, echo = TRUE)
```

# Introduction

## Airbnb Price Determinants in Europe

We want to work on Airbnb's dataset from kaggle.com. It provides information about hotel rooms in Europe.

Each major city has its dataset for weekends and weekdays
Variables included in the dataset: 
Host ID (Id)
The total price of listing (realSum)
Room type: private, shared, entire home, apt (room_type)
Whether or not a room is shared (room_shared)
Max number of people allowed in property (person_capacity)
Whether or not the host is superhost (host_is_superhost)
Whether or not it is multiple rooms (multi)
Whether for business or family use (biz)
Distance from the city center (dist)
Distance from nearest metro (metro_dist)
Latitude and longitude (lat long)
Guest satisfaction (guest_satisfaction_overall)
Cleanliness (cleanliness_rating)
The total quantity of bedrooms available among all properties for a single host (bedrooms)


Questions we can answer with the dataset:
Price Forecasting: use pricing, room type, and amenities to predict potential rental prices in the future.
Hotspots: use listing location in relation to business and tourism centers and correlate this with pricing to determine where Airbnb rentals would be most profitable
Customer Sentiment Analysis: analyze customer comments and satisfaction ratings to evaluate listing on overall customer experience and use it to optimize hosts' services to improve user satisfaction ratings.

How can this information be used:
Data can help travelers find accommodation that meets their needs without exceeding budget.
Can help hosts set competitive pricing and optimize listings to get more bookings.
Help investors evaluate the value of investing in real estate in different European cities based on pricing trends.

# Pre Processing and Cleaning the Data 

## Data loading

```{r}
# Set the relative directory path
my_dir <- "./archive"

# List all the files in the directory
files <- list.files(path = my_dir, full.names = TRUE)
```

### Combining the Data from all Files

```{r}
# Get a list of all the csv files in the directory
file_list <- list.files(path = my_dir, pattern = "*.csv", full.names = TRUE)

# Initialize an empty list to store the data frames
df_list <- list()

# Loop through each file and read it into a data frame
for (i in seq_along(file_list)) {
  df <- read.csv(file_list[i])
  
  # Add a new column with the city_day
  df$city_day <- basename(file_list[i])

  # Append the data frame to the list
  df_list[[i]] <- df
}

# Combine all the data frames into a single dataset
my_data <- bind_rows(df_list)

#Removing the .csv ext
my_data$city_day <- gsub("\\.csv", "", my_data$city_day)

# Print the first few rows of the data
head(my_data)
```


```{r}
print(unique(my_data[my_data$room_shared == my_data$room_private, ]$room_type)) # if the room is shared 
# summary(my_data)
```

```{r}
print(unique(my_data[my_data$room_private=='False', ]$room_type))

```

```{r}
print(unique(my_data[my_data$room_shared=='True', ]$room_type))
```

```{r}
print(unique(my_data[my_data$room_shared=='False', ]$room_type))
```

The room_shared and room_private information is already embedded in room_type. The variables are multi-collinear, so we can remove
room_shared and room_private.

### dropping columns of room_shared and room_private
```{r}
my_data = select(my_data, -c(room_shared, room_private))
head(my_data)

```

```{r, fig.height=10, fig.width=20}
ggplot(my_data, aes(x = realSum, fill = room_type, group = room_type)) + 
  geom_histogram(alpha = 0.5, nbins=20) + 
  theme(axis.title.x = element_text(size = 14),
        axis.title.y = element_text(size = 14))

ggplot(my_data_filtered, aes(x = realSum, fill = room_type, group = room_type)) + 
  geom_histogram(alpha = 0.5, nbins=20) + 
  theme(axis.title.x = element_text(size = 14),
        axis.title.y = element_text(size = 14))

ggplot(my_data_filtered, aes(x = realSum, fill = room_type, group = room_type)) + 
  geom_histogram(alpha = 0.5, nbins=20) + 
  theme(axis.title.x = element_text(size = 14),
        axis.title.y = element_text(size = 14)) + facet_wrap(~city_day)

```
The price of entire home/apt tend to be higher compared to other two categories. And the count of entire home /apt is also more.

```{r, fig.height=10, fig.width=20}
ggplot(my_data, aes(x = realSum, fill = person_capacity, group = person_capacity)) + 
  geom_histogram(alpha = 0.5, nbins=20) + 
  theme(axis.title.x = element_text(size = 14),
        axis.title.y = element_text(size = 14))

ggplot(my_data_filtered, aes(x = realSum, fill = person_capacity, group = person_capacity)) + 
  geom_histogram(alpha = 0.5, nbins=20) + 
  theme(axis.title.x = element_text(size = 14),
        axis.title.y = element_text(size = 14))

ggplot(my_data_filtered, aes(x = realSum, fill = person_capacity, group = person_capacity)) + 
  geom_histogram(alpha = 0.5, nbins=20) + 
  theme(axis.title.x = element_text(size = 14),
        axis.title.y = element_text(size = 14)) + facet_wrap(~city_day)

```
The overall price is distributed similarly across the spectrum irrespective of person_capacity. But for some cities like london, london_weekdays, lisbon the price is higher with person capacity. So, person capacity along with city will be an important variable for determining price.


```{r, fig.height=10, fig.width=20}
ggplot(my_data, aes(x = realSum, fill = host_is_superhost, group = host_is_superhost)) + 
  geom_histogram(alpha = 0.5, nbins=20) + 
  theme(axis.title.x = element_text(size = 14),
        axis.title.y = element_text(size = 14))

ggplot(my_data_filtered, aes(x = realSum, fill = host_is_superhost, group = host_is_superhost)) + 
  geom_histogram(alpha = 0.5, nbins=20) + 
  theme(axis.title.x = element_text(size = 14),
        axis.title.y = element_text(size = 14))

ggplot(my_data_filtered, aes(x = realSum, fill = host_is_superhost, group = host_is_superhost)) + 
  geom_histogram(alpha = 0.5, nbins=20) + 
  theme(axis.title.x = element_text(size = 14),
        axis.title.y = element_text(size = 14)) + facet_wrap(~city_day)

```
The prices are spread across all the spectrum irrespective of super_host or not.


```{r, fig.height=10, fig.width=20}
ggplot(my_data, aes(x = realSum, fill = multi, group = multi)) + 
  geom_histogram(alpha = 0.5, nbins=20) + 
  theme(axis.title.x = element_text(size = 14),
        axis.title.y = element_text(size = 14))

ggplot(my_data_filtered, aes(x = realSum, fill = multi, group = multi)) + 
  geom_histogram(alpha = 0.5, nbins=20) + 
  theme(axis.title.x = element_text(size = 14),
        axis.title.y = element_text(size = 14))

ggplot(my_data_filtered, aes(x = realSum, fill = multi, group = multi)) + 
  geom_histogram(alpha = 0.5, nbins=20) + 
  theme(axis.title.x = element_text(size = 14),
        axis.title.y = element_text(size = 14)) + facet_wrap(~city_day)

```
The prices are similar irrespctive of multi or not.

```{r, fig.height=10, fig.width=20}
ggplot(my_data, aes(x = realSum, fill = biz, group = biz)) + 
  geom_histogram(alpha = 0.5, nbins=20) + 
  theme(axis.title.x = element_text(size = 14),
        axis.title.y = element_text(size = 14))

ggplot(my_data_filtered, aes(x = realSum, fill = biz, group = biz)) + 
  geom_histogram(alpha = 0.5, nbins=20) + 
  theme(axis.title.x = element_text(size = 14),
        axis.title.y = element_text(size = 14))

ggplot(my_data_filtered, aes(x = realSum, fill = biz, group = biz)) + 
  geom_histogram(alpha = 0.5, nbins=20) + 
  theme(axis.title.x = element_text(size = 14),
        axis.title.y = element_text(size = 14)) + facet_wrap(~city_day)

```

The prices are similar irresptive of biz or not.


```{r, fig.height=10, fig.width=20}
ggplot(my_data, aes(x = realSum, fill = cleanliness_rating, group = cleanliness_rating)) + 
  geom_histogram(alpha = 0.5, nbins=20) + 
  theme(axis.title.x = element_text(size = 14),
        axis.title.y = element_text(size = 14))

ggplot(my_data_filtered, aes(x = realSum, fill = cleanliness_rating, group = cleanliness_rating)) + 
  geom_histogram(alpha = 0.5, nbins=20) + 
  theme(axis.title.x = element_text(size = 14),
        axis.title.y = element_text(size = 14))

ggplot(my_data_filtered, aes(x = realSum, fill = cleanliness_rating, group = cleanliness_rating)) + 
  geom_histogram(alpha = 0.5, nbins=20) + 
  theme(axis.title.x = element_text(size = 14),
        axis.title.y = element_text(size = 14)) + facet_wrap(~city_day)

```
The cleanliness rating doesn't really have an effect on price



### Scatterplot of Price vs Guest Satisfaction filtered by city

```{r, fig.height=10, fig.width=20}
ggplot(my_data_filtered, aes(x = realSum, y = guest_satisfaction_overall, color = city_day)) +
  geom_point() +
  xlab("Price") +
  ylab("Guest Satisfaction Overall") + 
  scale_color_discrete(name = "City-Day")

ggplot(my_data_filtered, aes(x = realSum, y = guest_satisfaction_overall, color = city_day)) +
  geom_point() +
  xlab("Price") +
  ylab("Guest Satisfaction Overall") + 
  scale_color_discrete(name = "City-Day") + facet_wrap(~city_day)
```
The plot depicts that there is no correlation of price with guest satisfaction, good satisfaction rate is found across all the prices.
In some cities like lonon, we can see a group of reviews with low guest satisfaction.



```{r, fig.height=10, fig.width=20}
ggplot(my_data, aes(x = realSum, fill = bedrooms, group = bedrooms)) + 
  geom_histogram(alpha = 0.6) + 
  theme(axis.title.x = element_text(size = 14),
        axis.title.y = element_text(size = 14))

ggplot(my_data_filtered, aes(x = realSum, fill = bedrooms, group = bedrooms)) + 
  geom_histogram(alpha = 0.6) + 
  theme(axis.title.x = element_text(size = 14),
        axis.title.y = element_text(size = 14))

ggplot(my_data_filtered, aes(x = realSum, fill = bedrooms, group = bedrooms)) + 
  geom_histogram(alpha = 0.6) + 
  theme(axis.title.x = element_text(size = 14),
        axis.title.y = element_text(size = 14)) + facet_wrap(~city_day)

```

```{r, fig.height=10, fig.width=20}
ggplot(my_data_filtered, aes(x = reorder(city_day, bedrooms, FUN=median), y = bedrooms, fill = city_day)) + 
  geom_boxplot() +
  coord_flip() + 
  theme(legend.key.height = unit(0.5, "cm"),
        legend.key.size = unit(1, "lines"))
```


```{r}
cor(as.numeric(factor(my_data$multi)), as.numeric(factor(my_data$biz)))
```



### Spilt Training and Testing Data

```{r}
set.seed(123456789)
my_data_train <- my_data[sample(nrow(my_data), 0.7*nrow(my_data)), ]
my_data_test <- my_data[setdiff(1:nrow(my_data), rownames(my_data_train)), ]
dim(my_data_train)
dim(my_data_test)
```

### Filtering out the Outliers from Data Out of IQR Ranges

```{r}
# Initialize an empty list to store the outliers
outliers_list <- list()

# Initialize an empty list to store the filtered data frames
df_list_filtered <- list()

# Loop through each file and read it into a data frame after removing outliers
for (i in seq_along(file_list)) {
  df_filtered <- read.csv(file_list[i])
  
  # Add a new column with the city_day
  df_filtered$city_day <- gsub("\\.csv", "", basename(file_list[i]))
  
  iqr_var1 <- IQR(df_filtered$realSum)

  # Calculate the upper and lower bounds for each variable
  upper_var1 <- quantile(df_filtered$realSum, 0.75) + 1.5*iqr_var1
  lower_var1 <- quantile(df_filtered$realSum, 0.25) - 1.5*iqr_var1


  # Filter the data based on the upper and lower bounds for each variable
  filtered_data <- filter(df_filtered, realSum > lower_var1 & realSum < upper_var1)


  # Append the filtered data frame to the list
  df_list_filtered[[i]] <- filtered_data
  
  # Get the rows that were removed while filtering
  outliers <- anti_join(df_filtered, filtered_data)
  
  # Append the outliers to the list
  outliers_list[[i]] <- outliers
}


# Combine all the filtered data frames into a single dataset
my_data_filtered <- bind_rows(df_list_filtered)

#Removing the .csv ext
my_data_filtered$city_day <- gsub("\\.csv", "", my_data_filtered$city_day)

# summary(my_data_filtered)

# Combine all the outliers into a single dataset
my_outliers <- bind_rows(outliers_list)

#Removing the .csv ext
my_outliers$city_day <- gsub("\\.csv", "", my_outliers$city_day)

summary(my_outliers)
```

# Exploratory Data Analysis

## Non Outlier Analysis

### Boxplot of Price Vs City

```{r}
ggplot(my_data_filtered, aes(x = reorder(city_day, realSum, FUN=median), y = realSum, fill = city_day)) + 
  geom_boxplot() +
  coord_flip() + 
  theme(legend.key.height = unit(0.5, "cm"),
        legend.key.size = unit(1, "lines"))

```

The highest prices in europe are found in amsterdam.

### Density plot of Price vs Room type

```{r}
ggplot(my_data_filtered, aes(x = realSum, group=room_type, fill=room_type, alpha=0.2)) + 
  geom_density() 
```

The prices of entire home are high comparitively





### Scatterplot of Prices in Rome w.r.t Latitude and Longitude during weekdays

```{r}
tema <- theme(plot.title = element_text(size = 23, hjust = .5),
              axis.text.x = element_text(size = 19, face = "bold"),
              axis.text.y = element_text(size = 19, face = "bold"),
              axis.title.x = element_text(size = 19),
              axis.title.y = element_text(size = 19),
              legend.text = element_text(colour="black", size=19, face="bold"),
              legend.background = element_rect(fill="#F5FFFA", size=0.5, linetype="dashed", 
                                  colour ="black"))

rome_data <- my_data_filtered %>% subset(city_day=='rome_weekdays')
  
ggplot(data = rome_data, mapping = aes(x = lat, y = lng)) +
    theme_minimal() +
    scale_fill_identity() +
    geom_point(mapping = aes(color = realSum), size = 3) +
    ggtitle("") +
    tema
```

This plot is within expectations of game theory, which suggests similar types of establishments (price and hospitality) tend be in clusters.

## Outlier Analysis

```{r}
ggplot() +
  geom_point(data = my_data_filtered, aes(x = realSum, y = metro_dist, color = "Filtered Data"), alpha = 0.4) +
  geom_point(data = my_outliers, aes(x = realSum, y = metro_dist, color = "Outliers"), alpha = 0.4) +
  scale_color_manual(values = c("Filtered Data" = "blue", "Outliers" = "red"))

ggplot() +
  geom_point(data = my_data_filtered, aes(x = realSum, y = metro_dist, color = "Filtered Data"), alpha = 0.4) +
  geom_point(data = my_outliers, aes(x = realSum, y = metro_dist, color = "Outliers"), alpha = 0.4) +
  scale_color_manual(values = c("Filtered Data" = "blue", "Outliers" = "red")) +
  facet_wrap(~city_day)
```

The pricey rooms are mostly near to the metro

```{r}
ggplot() +
  geom_point(data = my_data_filtered, aes(x = realSum, y = dist, color = "Filtered Data"), alpha = 0.4) +
  geom_point(data = my_outliers, aes(x = realSum, y = dist, color = "Outliers"), alpha = 0.4) +
  scale_color_manual(values = c("Filtered Data" = "blue", "Outliers" = "red"))

ggplot() +
  geom_point(data = my_data_filtered, aes(x = realSum, y = dist, color = "Filtered Data"), alpha = 0.4) +
  geom_point(data = my_outliers, aes(x = realSum, y = dist, color = "Outliers"), alpha = 0.4) +
  scale_color_manual(values = c("Filtered Data" = "blue", "Outliers" = "red")) +
  facet_wrap(~city_day)
```

The pricey rooms are mostly near to the centre of the city and the there is some correlation.

```{r}
ggplot() +
  geom_point(data = my_data_filtered, aes(x = realSum, y = attr_index, color = "Filtered Data"), alpha = 0.4) +
  geom_point(data = my_outliers, aes(x = realSum, y = attr_index, color = "Outliers"), alpha = 0.4) +
  scale_color_manual(values = c("Filtered Data" = "blue", "Outliers" = "red"))

ggplot() +
  geom_point(data = my_data_filtered, aes(x = realSum, y = attr_index, color = "Filtered Data"), alpha = 0.4) +
  geom_point(data = my_outliers, aes(x = realSum, y = attr_index, color = "Outliers"), alpha = 0.4) +
  scale_color_manual(values = c("Filtered Data" = "blue", "Outliers" = "red")) +
  facet_wrap(~city_day)

ggplot() +
  geom_point(data = my_data, aes(x = realSum, y = attr_index, color = "Filtered Data"), alpha = 0.4) 
```

The range of values falling b/w outliers and normal data is almost same . So there isn't a relationship b/w attr_index and realSum. 

```{r}
ggplot() +
  geom_point(data = my_data, aes(x = attr_index, y = attr_index_norm, color = "Filtered Data"), alpha = 0.4) +
  scale_color_manual(values = c("Filtered Data" = "blue", "Outliers" = "red"))

ggplot() +
  geom_point(data = my_data, aes(x = attr_index, y = attr_index_norm, color = "Filtered Data"), alpha = 0.4) +
  scale_color_manual(values = c("Filtered Data" = "blue", "Outliers" = "red")) +
  facet_wrap(~city_day)
```

attr_index and attr_index_norm are same, attr_index_norm is just normalized attr_index
 
```{r}
ggplot() +
  geom_point(data = my_data, aes(x = rest_index, y = rest_index_norm, color = "Filtered Data"), alpha = 0.4) +
  scale_color_manual(values = c("Filtered Data" = "blue", "Outliers" = "red"))

ggplot() +
  geom_point(data = my_data, aes(x = rest_index, y = rest_index_norm, color = "Filtered Data"), alpha = 0.4) +
  scale_color_manual(values = c("Filtered Data" = "blue", "Outliers" = "red")) +
  facet_wrap(~city_day)
```

rest_index and rest_index_norm are same, rest_index_norm is just normalized rest_index

```{r}
ggplot() +
  geom_point(data = my_data_filtered, aes(x = realSum, y = rest_index, color = "Filtered Data"), alpha = 0.4) +
  geom_point(data = my_outliers, aes(x = realSum, y = rest_index, color = "Outliers"), alpha = 0.4) +
  scale_color_manual(values = c("Filtered Data" = "blue", "Outliers" = "red"))

ggplot() +
  geom_point(data = my_data_filtered, aes(x = realSum, y = rest_index, color = "Filtered Data"), alpha = 0.4) +
  geom_point(data = my_outliers, aes(x = realSum, y = rest_index, color = "Outliers"), alpha = 0.4) +
  scale_color_manual(values = c("Filtered Data" = "blue", "Outliers" = "red")) +
  facet_wrap(~city_day)
```

There is no relationship outliers and rest_index

### Percentage of Outliers.

```{r}
# Create empty table
outliers_table <- data.frame(City_day = character(),
                             Data_Length = numeric(),
                             Percent_Outliers = numeric(),
                             stringsAsFactors = FALSE)

# Loop through city_data and fill in table
for (city_day in unique(my_data$city_day)){
  x = my_data[my_data$city_day == city_day, ]$realSum
  q1 <- quantile(x, 0.25)
  q3 <- quantile(x, 0.75)
  iqr <- IQR(x)
  upper_bound <- q3 + 1.5*iqr
  lower_bound <- q1 - 1.5*iqr
  x_no_outliers <- x[x >= lower_bound & x <= upper_bound]
  percent_outliers <- ((length(x) - length(x_no_outliers))/length(x))*100
  
  # Add row to table
  outliers_table <- rbind(outliers_table, 
                          data.frame(City_day = city_day, 
                                     Data_Length = length(x), 
                                     Percent_Outliers = percent_outliers))
}

# Format table using kable
kable(outliers_table, format = "markdown")
```

## Linear Regression

```{r}
M1 <- lm(realSum ~ . - realSum - city_day, data = my_data_train)
summary(M1)
```

### Checking for Multi Collinearity

```{r}
alias(M1)

# Calculate the VIFs
M2 <- lm(realSum ~ . - realSum - city_day - room_type, data = my_data_train)
vif(M2)
```

### Autoplots

```{r}
autoplot(M1) 
autoplot(M1, which = 4)
avPlots(M1)
```

### Cooks Distance

```{r}
cooksd <- cooks.distance(M1)
max(cooksd)
```

### Prediction Analysis

```{r}
my_data_test_spread <- my_data_test %>% spread_predictions(M1)
my_data_test_spread <- my_data_test %>% spread_residuals(M1)
mean(my_data_test_spread$M1)
```


## Lasso Regression

```{r}
x <- as.matrix(my_data[, c("room_type",
                           "room_shared",
                           "room_private",
                           "person_capacity",
                           "host_is_superhost",
                           "multi",
                           "biz",
                           "cleanliness_rating",
                           "guest_satisfaction_overall",
                           "bedrooms",
                           "dist",
                           "metro_dist",
                           "attr_index",
                           "attr_index_norm",
                           "rest_index",
                           "rest_index_norm",
                           "lng",
                           "lat",
                           "city_day")])
y <- my_data$realSum
lasso.fit <- glmnet(x, y, alpha = 1)
plot(lasso.fit, xvar = "lambda", label = TRUE)
```

# Model Selection

# Model Training

# Conclusion



```{r}
ggpairs(my_data[c("realSum","room_type","person_capacity","host_is_superhost","multi","biz","cleanliness_rating"
                  ,"guest_satisfaction_overall","bedrooms","dist","metro_dist"
                  ,"attr_index_norm","rest_index","city_day")], cardinality = 20, cardinality_threshold = 999)
```

```{r}
ggpairs(my_data[c("realSum","dist","metro_dist"
                  ,"attr_index_norm","rest_index_norm","city_day")], cardinality = 20, cardinality_threshold = 999)
```

```{r}
ggpairs(my_data_filtered[c("realSum","dist","metro_dist"
                  ,"attr_index_norm","rest_index_norm","city_day")], cardinality = 20, cardinality_threshold = 999)
```

```{r}
ggplot() +
  geom_point(data = my_data_filtered, aes(x = attr_index, y = rest_index, color = "Filtered Data"), alpha = 0.4) +
  geom_point(data = my_outliers, aes(x = attr_index, y = rest_index, color = "Outliers"), alpha = 0.4) +
  scale_color_manual(values = c("Filtered Data" = "blue", "Outliers" = "red"))

ggplot() +
  geom_point(data = my_data_filtered, aes(x = attr_index, y = rest_index, color = "Filtered Data"), alpha = 0.4) +
  geom_point(data = my_outliers, aes(x = attr_index, y = rest_index, color = "Outliers"), alpha = 0.4) +
  scale_color_manual(values = c("Filtered Data" = "blue", "Outliers" = "red")) +
  facet_wrap(~city_day)

ggplot() +
  geom_point(data = my_data, aes(x = attr_index, y = rest_index, color = "Filtered Data"), alpha = 0.4) +
  scale_color_manual(values = c("Filtered Data" = "blue", "Outliers" = "red"))

ggplot() +
  geom_point(data = my_data, aes(x = attr_index, y = rest_index, color = "Filtered Data"), alpha = 0.4) +
  scale_color_manual(values = c("Filtered Data" = "blue", "Outliers" = "red")) +
  facet_wrap(~city_day)
```

```{r}
cor(my_data$attr_index, my_data$rest_index)
```

```{r}
ggplot() +
  geom_point(data = my_data_filtered, aes(x = attr_index_norm, y = rest_index_norm, color = "Filtered Data"), alpha = 0.4) +
  geom_point(data = my_outliers, aes(x = attr_index_norm, y = rest_index_norm, color = "Outliers"), alpha = 0.4) +
  scale_color_manual(values = c("Filtered Data" = "blue", "Outliers" = "red"))

ggplot() +
  geom_point(data = my_data_filtered, aes(x = attr_index_norm, y = rest_index_norm, color = "Filtered Data"), alpha = 0.4) +
  geom_point(data = my_outliers, aes(x = attr_index_norm, y = rest_index_norm, color = "Outliers"), alpha = 0.4) +
  scale_color_manual(values = c("Filtered Data" = "blue", "Outliers" = "red")) +
  facet_wrap(~city_day)

ggplot() +
  geom_point(data = my_data, aes(x = attr_index_norm, y = rest_index_norm, color = "Filtered Data"), alpha = 0.4) +
  scale_color_manual(values = c("Filtered Data" = "blue", "Outliers" = "red"))

ggplot() +
  geom_point(data = my_data, aes(x = attr_index_norm, y = rest_index_norm, color = "Filtered Data"), alpha = 0.4) +
  scale_color_manual(values = c("Filtered Data" = "blue", "Outliers" = "red")) +
  facet_wrap(~city_day)
```

```{r}
cor(my_data$attr_index_norm, my_data$rest_index_norm)
```



```{r}
my_data[c("realSum","room_type","person_capacity","host_is_superhost","multi","biz","cleanliness_rating"
                  ,"guest_satisfaction_overall","bedrooms","dist","metro_dist"
                  ,"attr_index_norm","rest_index","city_day")]
```





